---
title: "covertest() - a simple tool for training the estimation of vegetation cover"
author: "Roman M. Link"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
```

## Motivation
After reading an interesting question by [Ekaterina Petrunenko](https://www.researchgate.net/profile/Ekaterina_Petrunenko) on [ResearchGate][https://www.researchgate.net/post/Software_for_training_accurate_eye], I realized that I (as well as probably most other field biologists) often struggled with precise vegetation coverage estimates because in nature there is no easy way to practice it.

I therefore wrote a small little script (`covertest.R`) that generates samples of randomly generated vegetation cover, allowing you to improve your estimation skills.

## How it works
To achieve this, I draw a sample from a multivariate normal distribution with a Gaussian spatial autocorrelation function with correlation parameter `delta`. I then randomly sample a proportion from a specified range (0 - 100 % unless `range` is changed from its original values), and classify the raster into two groups based on the threshold specified by the randomly sampled proportion. 

This raster is then printed, and the original and classified raster, the true coverage percentage and the autocorrelation parameter are then silently returned. A print function is defined that allows to easily check the true coverage percentage, so it is easy to first look at the plot, make a guess, and then check the true value.

## Example
The functions in this script rely on the functionalities of the `raster` package, so do not forget to load (and if necessary install) it before starting.

```{r, echo=TRUE}
# install raster package if necessary
# install.packages("raster")
library(raster)
```

Then, simply copy the `covertest.R` file from this GitHub project into your project folder and load it with `source()`.
```{r, echo=TRUE}
source("covertest.R")
```


Now you just have to run the covertest function to get a plot of random "vegetation cover".
```{r, echo=TRUE}
# random seed for replicability
set.seed(12345)

# create simulated raster
cover <- covertest(delta = 5, n = 30, range = c(0, 100))

```


To get the true coverage percentage, simply print the generated object.
```{r, echo=TRUE}
# print to get output 
cover
```

To practice estimating the vegetation under different conditions, play around with the delta parameter. Large values of `delta` produce large areas of contiguous coverage:
```{r, echo=TRUE}
set.seed(56789)
cover1 <- covertest(delta = 20, n = 30, range = c(0, 100))
cover1
```

Low values of `delta` lead to more small-scale variability in the "vegetation cover".
```{r, echo=TRUE}
set.seed(12321)
cover2 <- covertest(delta = 1, n = 30, range = c(0, 100))
cover2
```

Changing `n` changes the number of cells in the analyzed n x n matrix. Be careful, high values of `n` lead to large computation times!

```{r, echo=TRUE}
set.seed(23456)
cover2 <- covertest(delta = 10, n = 50, range = c(0, 100))
cover2
```



The `range` parameter can be used to set the range of coverage probabilities. If you use the same value for the lower and upper bound, you can get simulations for a specific coverage proportion, say 37.6% (up to rounding errors caused by the finite number of cells):
```{r, echo=TRUE}
set.seed(12321)
cover2 <- covertest(delta = 2.5, n = 30, range = c(37.6, 37.6))
cover2
```

## Last words
Please enjoy my script, but note that it is just the outcome of a friday afternoon procrastination session and as such comes without any warranty whatsoever...